---
title: "Veeraraghavan etal 2017"
author: "Harini Veeraraghavan and James D. Brenton"
date: "18 March 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, echo=FALSE, warning= FALSE}
require(tidyverse)
require(knitr)
require(broom)
require(survival)
require(survminer)
require(plyr)
require(ggplot2)

# options for output
options(scipen=1, digits=2)
# Variants of mean, median etc with na.rm = T
min_    <- function(...) min(..., na.rm=T)
max_    <- function(...) max(..., na.rm=T)
mean_   <- function(...) mean(..., na.rm=T)
median_ <- function(...) median(..., na.rm=T)
Q1_     <- function(...) quantile(..., probs=0.25, na.rm = TRUE)
Q3_     <- function(...) quantile(..., probs=0.75, na.rm = TRUE)
sum_    <- function(...) sum(..., na.rm=T)
# function to return counts of samples
n_fun <- function(x, y){ # y is position for labels
  return(data.frame(y, label = paste0("n = ", length(x))))
}
```

Load in clinical data.

```{r, echo = FALSE, results="hide", warning=FALSE}
clin.data <- read_csv("2017-Veeraraghavan-data-HV.csv")
```

## Simple data exploration

Total number of patients

```{r patients_total, echo = FALSE}
clin.data %>%
  distinct(tcga.id) %>% tally()
```


####Were there differences in the number of sites (ROI) between the cohorts?

```{r, echo = FALSE, message=FALSE}
clin.data %>%
    group_by(institution.bin) %>%
    summarize_each(funs(
    min_,
    Q1_,
    median_,
    Q3_,
    max_
    ), 
    sites) 
```

Plot the ROI counts.

```{r, echo = FALSE, warning=FALSE}
clin.data %>%
  ggplot(aes(x = institution.bin, y = sites)) + geom_boxplot() +
    stat_summary(fun.data = n_fun, fun.args = 15, geom = "text")

print(wilcox.test(clin.data$sites[clin.data$institution.bin=="MSKCC"], 
                  clin.data$sites[clin.data$institution.bin != "MSKCC"]))
```


####Differences in OS in the two cohorts? (TCIA and MSKCC)

```{r, echo=FALSE, warning=FALSE}
clin.data %>%
    group_by(institution.bin) %>%
    summarize_each(funs(
    min_,
    Q1_,
    median_,
    Q3_,
    max_
    ), 
    os.mos) 
```


```{r, echo=FALSE}
clin.data %>% 
  group_by(institution.bin) %>% 
  do(glance(survfit(Surv(time=os.mos, event=dead=="1") ~ 1, data=.)))

```

#####Plot comparison of OS.

```{r, echo=FALSE}
ggsurvplot(
  survfit(Surv(time=os.mos, event=dead=="1") ~ institution.bin, data=clin.data),
  data = clin.data, risk.table = TRUE, conf.int = TRUE, pval = TRUE)
```


####Differences in PFS in the two cohorts? (TCIA and MSKCC)

```{r, echo=FALSE, warning=FALSE}
clin.data %>%
    group_by(institution.bin) %>%
    summarize_each(funs(
    min_,
    Q1_,
    median_,
    Q3_,
    max_
    ), 
    pfs.mos) 
```

```{r, echo=FALSE}
clin.data %>% 
  group_by(institution.bin) %>% 
  do(glance(survfit(Surv(time=pfs.mos, event=recurrence=="1") ~ 1, data=.)))
```

#####Plot comparison of OS.

```{r, echo=FALSE}
ggsurvplot(
  survfit(Surv(time=pfs.mos, event=recurrence=="1") ~ institution.bin, data=clin.data),
  data = clin.data, risk.table = TRUE, conf.int = TRUE, pval = TRUE)
```


####Differences in follow ups

```{r, echo=FALSE, warning=FALSE}
clin.data %>%
    group_by(institution.bin) %>%
    summarize_each(funs(
    min_,
    Q1_,
    median_,
    Q3_,
    max_
    ), 
    Follow.up.mos) 
 
```

#####Plot comparison of follow ups
```{r, echo=FALSE, warning=FALSE}

clin.data %>%
  ggplot(aes(x = institution.bin, y = Follow.up.mos)) + geom_boxplot() +
    stat_summary(fun.data = n_fun, fun.args = 15, geom = "text")

print(wilcox.test(clin.data$sites[clin.data$institution.bin=="MSKCC"], 
                  clin.data$sites[clin.data$institution.bin != "MSKCC"]))
  
```

#### Differences in total tumor volume between the datasets
```{r, echo=FALSE, warning=FALSE}
clin.data %>%
    group_by(institution.bin) %>%
    summarize_each(funs(
    min_,
    Q1_,
    median_,
    Q3_,
    max_
    ), 
    volume.cc) 
```

Plot the volume distributions.

```{r, echo = FALSE}
clin.data %>%
  ggplot(aes(x = institution.bin, y = volume.cc)) + geom_boxplot() +
    stat_summary(fun.data = n_fun, fun.args = 2586, geom = "text")
```

Non-parametric test for volume distributions.

```{r, echo = FALSE}
clin.data %>% 
  do(tidy(wilcox.test(volume.cc ~ institution.bin, data = .)))
```

***


####Differences in the computed texture measures between the datasets
#####Energy

```{r, echo=FALSE, warning=FALSE}
  clin.data %>%
    group_by(institution.bin) %>%
    summarize_each(funs(
    min_,
    Q1_,
    median_,
    Q3_,
    max_
    ), 
    energy)

```

Plot the energy distribution

```{r, echo = FALSE}
clin.data %>%
  ggplot(aes(x = institution.bin, y = energy)) + geom_boxplot() +
    stat_summary(fun.data = n_fun, fun.args = 256, geom = "text")
```

Non-parametric test for energy distributions.

```{r, echo = FALSE}
clin.data %>% 
  do(tidy(wilcox.test(energy ~ institution.bin, data = .)))
```

#####Entropy

```{r, echo=FALSE, warning=FALSE}
  clin.data %>%
    group_by(institution.bin) %>%
    summarize_each(funs(
    min_,
    Q1_,
    median_,
    Q3_,
    max_
    ), 
    entropy)

```

Plot the energy distribution

```{r, echo = FALSE}
clin.data %>%
  ggplot(aes(x = institution.bin, y = entropy)) + geom_boxplot() +
    stat_summary(fun.data = n_fun, fun.args = 256, geom = "text")
```

Non-parametric test for entropy distributions.

```{r, echo = FALSE}
clin.data %>% 
  do(tidy(wilcox.test(entropy ~ institution.bin, data = .)))
```


#####Contrast

```{r, echo=FALSE, warning=FALSE}
  clin.data %>%
    group_by(institution.bin) %>%
    summarize_each(funs(
    min_,
    Q1_,
    median_,
    Q3_,
    max_
    ), 
    contrast)

```

Plot the contrast distribution

```{r, echo = FALSE}
clin.data %>%
  ggplot(aes(x = institution.bin, y = contrast)) + geom_boxplot() +
    stat_summary(fun.data = n_fun, fun.args = 256, geom = "text")
```

Non-parametric test for contrast distributions.

```{r, echo = FALSE}
clin.data %>% 
  do(tidy(wilcox.test(contrast ~ institution.bin, data = .)))
```

#####Homogeneity

```{r, echo=FALSE, warning=FALSE}
  clin.data %>%
    group_by(institution.bin) %>%
    summarize_each(funs(
    min_,
    Q1_,
    median_,
    Q3_,
    max_
    ), 
    homogeneity)

```

Plot the homogeneity distribution

```{r, echo = FALSE}
clin.data %>%
  ggplot(aes(x = institution.bin, y = homogeneity)) + geom_boxplot() +
    stat_summary(fun.data = n_fun, fun.args = 256, geom = "text")
```

Non-parametric test for homogeneity distributions.

```{r, echo = FALSE}
clin.data %>% 
  do(tidy(wilcox.test(homogeneity ~ institution.bin, data = .)))
```

#####cSE

```{r, echo=FALSE, warning=FALSE}
  clin.data %>%
    group_by(institution.bin) %>%
    summarize_each(funs(
    min_,
    Q1_,
    median_,
    Q3_,
    max_
    ), 
    c.se)

```

Plot the cSE distribution

```{r, echo = FALSE}
clin.data %>%
  ggplot(aes(x = institution.bin, y = c.se)) + geom_boxplot() +
    stat_summary(fun.data = n_fun, fun.args = 6, geom = "text")
```

Non-parametric test for cSE distributions.

```{r, echo = FALSE}
clin.data %>% 
  do(tidy(wilcox.test(c.se ~ institution.bin, data = .)))
```

#####cluDev

```{r, echo=FALSE, warning=FALSE}
  clin.data %>%
    group_by(institution.bin) %>%
    summarize_each(funs(
    min_,
    Q1_,
    median_,
    Q3_,
    max_
    ), 
    clu.dev)

```

Plot the cluDev distribution

```{r, echo = FALSE}
clin.data %>%
  ggplot(aes(x = institution.bin, y = clu.dev)) + geom_boxplot() +
    stat_summary(fun.data = n_fun, fun.args = 25, geom = "text")
```

Non-parametric test for cluDev distributions.

```{r, echo = FALSE}
clin.data %>% 
  do(tidy(wilcox.test(clu.dev ~ institution.bin, data = .)))
```

#####cluDiss

```{r, echo=FALSE, warning=FALSE}
  clin.data %>%
    group_by(institution.bin) %>%
    summarize_each(funs(
    min_,
    Q1_,
    median_,
    Q3_,
    max_
    ), 
    clu.diss)

```


Plot the cluDiss distribution

```{r, echo = FALSE}
clin.data %>%
  ggplot(aes(x = institution.bin, y = clu.diss)) + geom_boxplot() +
    stat_summary(fun.data = n_fun, fun.args = 30902, geom = "text")
```

Non-parametric test for cluDiss distributions.

```{r, echo = FALSE}
clin.data %>% 
  do(tidy(wilcox.test(clu.diss ~ institution.bin, data = .)))
```

***
##### Spearman rank correlation between texture measures and tumour burden

cSE vs. tumour volume

```{r, echo=FALSE, warning=FALSE}
    clin.data %>%
      do(tidy(cor.test(clin.data$c.se, clin.data$volume.cc, method="spearman"))) %>%
  select(estimate, p.value)
```

cluDiss vs. tumour volume

```{r, echo=FALSE, warning=FALSE}
    clin.data %>%
      do(tidy(cor.test(clin.data$clu.diss, clin.data$volume.cc, method="spearman"))) %>%
  select(estimate, p.value)
```

cluDev vs. tumour volume

```{r, echo=FALSE, warning=FALSE}

   clin.data %>%
      do(tidy(cor.test(clin.data$clu.dev, clin.data$volume.cc, method="spearman"))) %>%
  select(estimate, p.value)

```

Energy vs. tumour volume

```{r, echo=FALSE, warning=FALSE}
      
    clin.data %>%
      do(tidy(cor.test(clin.data$energy, clin.data$volume.cc, method="spearman"))) %>%
  select(estimate, p.value)

```

Entropy vs. tumour volume

```{r, echo=FALSE, warning=FALSE}
  
    clin.data %>%
      do(tidy(cor.test(clin.data$entropy, clin.data$volume.cc, method="spearman"))) %>%
  select(estimate, p.value)
```

Contrast vs. tumour volume

```{r, echo=FALSE, warning=FALSE}

    clin.data %>%
      do(tidy(cor.test(clin.data$contrast, clin.data$volume.cc, method="spearman"))) %>%
  select(estimate, p.value)

```

Homogeneity vs. tumour volume

```{r, echo=FALSE, warning=FALSE}

    clin.data %>%
      do(tidy(cor.test(clin.data$homogeneity, clin.data$volume.cc, method="spearman"))) %>%
  select(estimate, p.value)

```

Correlation with number of sites 

cSE vs. sites

```{r, echo=FALSE, warning=FALSE}
    clin.data %>%
      do(tidy(cor.test(clin.data$c.se, clin.data$sites, method="spearman"))) %>%
  select(estimate, p.value)
```

cluDiss vs. sites

```{r, echo=FALSE, warning=FALSE}

    clin.data %>%
      do(tidy(cor.test(clin.data$clu.diss, clin.data$sites, method="spearman"))) %>%
  select(estimate, p.value)
```

cluDev vs. sites

```{r, echo=FALSE, warning=FALSE}

   clin.data %>%
      do(tidy(cor.test(clin.data$clu.dev, clin.data$sites, method="spearman"))) %>%
  select(estimate, p.value)
```

Energy vs. sites

```{r, echo=FALSE, warning=FALSE}

    clin.data %>%
      do(tidy(cor.test(clin.data$energy, clin.data$sites, method="spearman"))) %>%
  select(estimate, p.value)

```
Entropy vs. sites

```{r, echo=FALSE, warning=FALSE}

    clin.data %>%
      do(tidy(cor.test(clin.data$entropy, clin.data$sites, method="spearman"))) %>%
  select(estimate, p.value)
    
```

Contrast vs. sites

```{r, echo=FALSE, warning=FALSE}

    clin.data %>%
      do(tidy(cor.test(clin.data$contrast, clin.data$sites, method="spearman"))) %>%
  select(estimate, p.value)

```

Homogeneity vs. sites

```{r, echo=FALSE, warning=FALSE}

    clin.data %>%
      do(tidy(cor.test(clin.data$homogeneity, clin.data$sites, method="spearman"))) %>%
  select(estimate, p.value)

```

*** 

##Volume and association to outcomes

##### Cox regression to determine if volume was associated with survival in the two datasets

Check if volume was associated with PFS

```{r, echo=FALSE, warning=FALSE}
  clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=pfs.mos, event=recurrence=="1")~volume.cc+age + as.factor(stage), data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)

```

Check if volume was associated with OS

```{r, echo=FALSE, warning=FALSE}
 clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=os.mos, event=dead=="1")~volume.cc+age + as.factor(stage), data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)

```
  

##### Non-parametric two-sided analysis for volume vs. platinum status

```{r, echo=FALSE, warning=FALSE}
clin.data %>% 
  group_by(institution.bin) %>%
    do(tidy(wilcox.test(volume.cc ~ platinum.resistant, data =.))) %>%
  select(institution.bin, statistic, p.value) %>%
  arrange(institution.bin, p.value)


```

***


##ISTH measures and association to outcomes

####Cox hazard regression for continuous ISTH measures with PFS in the two datasets

Check if cSE was associated with PFS 

```{r, echo=FALSE, warning=FALSE}
    clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=pfs.mos, event=recurrence=="1")~c.se+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)
```

Check if cSE was associated with OS

```{r, echo=FALSE, warning=FALSE}
 clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=os.mos, event=dead=="1")~c.se+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)

```

Check if cluDev was associated with PFS 

```{r, echo=FALSE, warning=FALSE}
    clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=pfs.mos, event=recurrence=="1")~clu.dev+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)
```

Check if cluDev was associated with OS

```{r, echo=FALSE, warning=FALSE}
 clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=os.mos, event=dead=="1")~clu.dev+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)

```

Check if cluDiss was associated with PFS 

```{r, echo=FALSE, warning=FALSE}
    clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=pfs.mos, event=recurrence=="1")~clu.diss+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)
```

Check if cluDiss was associated with OS

```{r, echo=FALSE, warning=FALSE}
 clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=os.mos, event=dead=="1")~clu.diss+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)

```

#### Non-parameteric tests to test association of ISTH measures with platinum status

Was cSE associated with Platinum sensitivity

```{r, echo=FALSE, warning=FALSE}
clin.data %>% 
  group_by(institution.bin) %>%
    do(tidy(wilcox.test(c.se ~ platinum.resistant, data =.))) %>%
  select(institution.bin, statistic, p.value) %>%
  arrange(institution.bin, p.value)

```

Was cluDev associated with Platinum sensitivity

```{r, echo=FALSE, warning=FALSE}
clin.data %>% 
  group_by(institution.bin) %>%
    do(tidy(wilcox.test(clu.dev ~ platinum.resistant, data =.))) %>%
  select(institution.bin, statistic, p.value) %>%
  arrange(institution.bin, p.value)

```

Was cluDiss associated with Platinum sensitivity

```{r, echo=FALSE, warning=FALSE}
clin.data %>% 
  group_by(institution.bin) %>%
    do(tidy(wilcox.test(clu.diss ~ platinum.resistant, data =.))) %>%
  select(institution.bin, statistic, p.value) %>%
  arrange(institution.bin, p.value)

```

#####Distribution of ISTH measures across cluster groups -- category boundaries

cSE

```{r, echo=FALSE, warning=FALSE}

  clin.data%>%
    group_by(institution.bin, ClusterID.ISTH) %>%
      summarize_each(funs(
        min_,
        Q1_,
        median_,
        Q3_,
        max_
      ), c.se)

```

cluDev

```{r, echo=FALSE, warning=FALSE}

 clin.data%>%
    group_by(institution.bin, ClusterID.ISTH) %>%
      summarize_each(funs(
        min_,
        Q1_,
        median_,
        Q3_,
        max_
      ), clu.dev)
```

cluDiss

```{r, echo=FALSE, warning=FALSE}

  clin.data%>%
    group_by(institution.bin, ClusterID.ISTH) %>%
      summarize_each(funs(
        min_,
        Q1_,
        median_,
        Q3_,
        max_
      ), clu.diss)

```


####Association between high, ultra-high vs. low, medium ISTH clusters and survival

Was low,medium cluster heterogeneity associated with PFS?

```{r, echo=FALSE, warning=FALSE}
    clin.data$clusterISTH <- (clin.data$ClusterID.ISTH == "Low" | clin.data$ClusterID.ISTH=="Medium")
    slo <- clin.data$clusterISTH
    shi <- !clin.data$clusterISTH
    clin.data$clusterISTH[slo] <- paste("Low")
    clin.data$clusterISTH[shi] <- paste("High")
    clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=pfs.mos, event=recurrence=="1")~clusterISTH+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)
```

Was low,medium cluster heterogeneity associated with OS?

```{r, echo=FALSE, warning=FALSE}
    clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=os.mos, event=dead=="1")~clusterISTH+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)
```

#### Distribution of clinical characteristics in the clustered groups for the two datasets

Number of ROIs

```{r, echo=FALSE, warning= FALSE}
  clin.data %>%
    group_by(institution.bin,ClusterID.ISTH) %>%
        summarize_each(funs(
    min_,
    Q1_,
    median_,
    Q3_,
    max_,
    sum_), sites)
```

Volume 


```{r, echo=FALSE, warning= FALSE}
  clin.data %>%
    group_by(institution.bin,ClusterID.ISTH) %>%
        summarize_each(funs(
    min_,
    Q1_,
    median_,
    Q3_,
    max_,
    sum_), volume.cc)
```

Number of cases

```{r, echo=FALSE, warning=FALSE}
  
  clin.data %>%
    group_by(institution.bin, ClusterID.ISTH) %>%
      do(data.frame(nrow=nrow(.)))
```

PFS 

```{r, echo=FALSE, warning=FALSE}

  clin.data%>%
    group_by(institution.bin,ClusterID.ISTH) %>%
      summarize_each(funs(
        min_,
        Q1_,
        median_,
        Q3_,
        max_
      ), pfs.mos)
```

OS 

```{r, echo=FALSE, warning=FALSE}

  clin.data%>%
    group_by(institution.bin,ClusterID.ISTH) %>%
      summarize_each(funs(
        min_,
        Q1_,
        median_,
        Q3_,
        max_
      ), os.mos)
```

Resection status

```{r, echo=FALSE, warning=FALSE}
  
  t <- clin.data%>%
    group_by(institution.bin,ClusterID.ISTH,residual.disease) %>%
       do(data.frame(nrow=nrow(.)))
  t <- na.omit(t)
  print(t,n=50)
```



Platinum resistant

```{r, echo=FALSE, warning=FALSE}

  t <- clin.data%>%
    group_by(institution.bin,ClusterID.ISTH,platinum.resistant) %>%
       do(data.frame(nrow=nrow(.)))
  print(t,n=21)
  t <- na.omit(t)
```

Fisher exact test for MSKCC clusters ISTH vs. platinum resistance

```{r, echo=FALSE, warning=FALSE}

  tidy(fisher.test(t(matrix(as.matrix.data.frame(t[1:8,4]), 
                            nrow=2,ncol=4, 
                            dimnames = list(c("Sensitive", "Resistant"), c("High", "Low", "Medium", "Ultra-High")))), 
                   alternative="two.sided"))

```

Fisher exact test for TCIA clusters ISTH vs. platinum resistance
```{r, echo=FALSE, warning=FALSE}
    tidy(fisher.test(t(matrix(as.matrix.data.frame(t[9:17,4]), 
                            nrow=2,ncol=4, 
                            dimnames = list(c("Sensitive", "Resistant"), c("High", "Low", "Medium", "Ultra-High")))), 
                   alternative="two.sided"))
```



#### Prevalence of Platinum resistance in clusters

Platinum resistance/sensitivity in low,medium vs. high/ultra-high ISTH clusters

```{r, echo=FALSE, warning=FALSE}
  t <- clin.data%>%  
    group_by(institution.bin, clusterISTH, platinum.resistant) %>%
      do(data.frame(nrow=nrow(.)))
  print(t)
  
  
```

Fisher exact test for MSKCC clusters ISTH vs. platinum resistance

```{r, echo=FALSE, warning=FALSE}

  platinumByClusters <- matrix(c(21, 14, 2, 7), nrow=2, dimnames = list(c("Sensitive", "Resistant"), c("Low", "High")))

  tidy(fisher.test(platinumByClusters, alternative = "two.sided")) %>%
    select(estimate, p.value, conf.low, conf.high)
```

Fisher exact test for TCIA clusters ISTH vs. platinum resistance
```{r, echo=FALSE, warning=FALSE}
   platinumByClusters <- matrix(c(7, 10, 3, 4), nrow=2, dimnames = list(c("Sensitive", "Resistant"), 
                                                                       c("Low", "High")))
  tidy(fisher.test(platinumByClusters, alternative="two.sided")) %>%
    select(estimate, p.value, conf.low, conf.high)
```

#### Prevalence of CCNE1 amplifications in the datasets

```{r, echo=FALSE, warning=FALSE}
    t <- clin.data%>%  
    group_by(institution.bin, clusterISTH, CCNE1.amp) %>%
      do(data.frame(nrow=nrow(.)))
  print(t)
  t <- na.omit(t)
```

Fisher exact test for clusters vs. CCNE1: MSKCC

```{r, echo=FALSE, warning=FALSE}
  tidy(fisher.test(t(matrix(as.matrix.data.frame(t[1:4,4]), 
                            nrow=2,ncol=2, 
                            dimnames = list(c("NoAmp", "Amp"), c("High", "Low")))), 
                   alternative="two.sided"))
```

Fisher exact test for clusters vs. CCNE1: TCIA

```{r, echo=FALSE, warning=FALSE}
  tidy(fisher.test(t(matrix(as.matrix.data.frame(t[5:8,4]), 
                            nrow=2,ncol=2, 
                            dimnames = list(c("NoAmp", "Amp"), c("High", "Low")))), 
                   alternative="two.sided"))
```

Resection status

```{r, echo=FALSE, warning=FALSE}
  
  tdata <- clin.data%>%
    group_by(institution.bin, clusterISTH,residual.disease) %>%
       do(data.frame(nrow=nrow(.)))
  tdata <- na.omit(tdata)
  print(tdata,n=50)
```

Fisher exact test for clusters vs. resection: MSKCC

```{r, echo=FALSE, warning=FALSE}

  print(t(matrix(as.matrix(tdata[1:6,4]), nrow=3, ncol=2, dimnames = list(c("Complete", "Optimal", "Suboptimal"), c("High", "Low")))))

  tidy(fisher.test(t(matrix(as.matrix(tdata[1:6,4]), nrow=3, ncol=2, dimnames = list(c("Complete", "Optimal", "Suboptimal"), c("High", "Low")))), alternative = "two.sided"))

```

Fisher exact test for clusters vs. resection: TCIA

```{r, echo=FALSE, warning=FALSE}

  print(t(matrix(as.matrix(tdata[7:13,4]), nrow=3, ncol=2, dimnames = list(c("Complete", "Optimal", "Suboptimal"), c("High", "Low")))))

  tidy(fisher.test(t(matrix(as.matrix(tdata[7:13,4]), nrow=3, ncol=2, dimnames = list(c("Complete", "Optimal", "Suboptimal"), c("High", "Low")))), alternative = "two.sided"))

```


***

##Haralick textures and association to outcomes
####Cox hazard regression for continuous Haralick texture measures with PFS in the two datasets

Check if Energy was associated with PFS 

```{r, echo=FALSE, warning=FALSE}
    clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=pfs.mos, event=recurrence=="1")~energy+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)
```

Check if Energy was associated with OS

```{r, echo=FALSE, warning=FALSE}
 clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=os.mos, event=dead=="1")~energy+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)

```

Check if Entropy was associated with PFS 

```{r, echo=FALSE, warning=FALSE}
    clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=pfs.mos, event=recurrence=="1")~entropy+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)
```

Check if Entropy was associated with OS

```{r, echo=FALSE, warning=FALSE}
 clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=os.mos, event=dead=="1")~entropy+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)

```

Check if Contrast was associated with PFS 

```{r, echo=FALSE, warning=FALSE}
    clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=pfs.mos, event=recurrence=="1")~contrast+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)
```

Check if Contrast was associated with OS

```{r, echo=FALSE, warning=FALSE}
 clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=os.mos, event=dead=="1")~contrast+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)

```

Check if Homogeneity was associated with PFS 

```{r, echo=FALSE, warning=FALSE}
    clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=pfs.mos, event=recurrence=="1")~homogeneity+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)
```

Check if Homogeneity was associated with OS

```{r, echo=FALSE, warning=FALSE}
 clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=os.mos, event=dead=="1")~homogeneity+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)

```

#### Non-parameteric tests to test association of Haralick textures with platinum status


```{r, echo=FALSE, warning=FALSE}
clin.data %>% 
  group_by(institution.bin) %>%
    do(tidy(wilcox.test(energy ~ platinum.resistant, data =.))) %>%
  select(institution.bin, statistic, p.value) %>%
  arrange(institution.bin, p.value)

```

Was entropy associated with Platinum sensitivity

```{r, echo=FALSE, warning=FALSE}
clin.data %>% 
  group_by(institution.bin) %>%
    do(tidy(wilcox.test(entropy ~ platinum.resistant, data =.))) %>%
  select(institution.bin, statistic, p.value) %>%
  arrange(institution.bin, p.value)

```

Was contrast associated with Platinum sensitivity

```{r, echo=FALSE, warning=FALSE}
clin.data %>% 
  group_by(institution.bin) %>%
    do(tidy(wilcox.test(contrast ~ platinum.resistant, data =.))) %>%
  select(institution.bin, statistic, p.value) %>%
  arrange(institution.bin, p.value)

```


Was homogeneity associated with Platinum sensitivity

```{r, echo=FALSE, warning=FALSE}
clin.data %>% 
  group_by(institution.bin) %>%
    do(tidy(wilcox.test(homogeneity ~ platinum.resistant, data =.))) %>%
  select(institution.bin, statistic, p.value) %>%
  arrange(institution.bin, p.value)

```



###### Category boundaries for Haralick clusters

Energy

```{r, echo=FALSE, warning=FALSE}

  clin.data%>%
    group_by(institution.bin, ClusterID.Haralick) %>%
      summarize_each(funs(
        min_,
        Q1_,
        median_,
        Q3_,
        max_
      ), energy)

```

Entropy

```{r, echo=FALSE, warning=FALSE}

  clin.data%>%
    group_by(institution.bin, ClusterID.Haralick) %>%
      summarize_each(funs(
        min_,
        Q1_,
        median_,
        Q3_,
        max_
      ), entropy)

```

Contrast

```{r, echo=FALSE, warning=FALSE}

  clin.data%>%
    group_by(institution.bin, ClusterID.Haralick) %>%
      summarize_each(funs(
        min_,
        Q1_,
        median_,
        Q3_,
        max_
      ), contrast)

```

Homogeneity

```{r, echo=FALSE, warning=FALSE}

  clin.data%>%
    group_by(institution.bin, ClusterID.Haralick) %>%
      summarize_each(funs(
        min_,
        Q1_,
        median_,
        Q3_,
        max_
      ), homogeneity)

```





####Association between Haralick clusters and survival

Haralick clusters and association with PFS?

```{r, echo=FALSE, warning=FALSE}

    clin.data$Hcluster <- clin.data$ClusterID.Haralick
    clin.data$Hcluster[clin.data$Hcluster=="Low"] <- paste("aLow")
    clin.data$Hcluster[clin.data$Hcluster=="Medium"] <- paste("bMedium")
    clin.data$Hcluster[clin.data$Hcluster=="High"] <- paste("cHigh")
      

    clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=pfs.mos, event=recurrence=="1")~Hcluster+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)
```

Haralick clusters and association with OS?

```{r, echo=FALSE, warning=FALSE}
    clin.data %>%
    group_by(institution.bin) %>%
      do(tidy(coxph(Surv(time=os.mos, event=dead=="1")~Hcluster+age+as.factor(stage)+volume.cc, data=.), exponentiate=TRUE)) %>%
    select(institution.bin, term, estimate, p.value, conf.low, conf.high) %>%
  arrange(institution.bin, p.value)
```

#### Distribution of clinical characteristics in the clustered groups for the two datasets

Number of ROIs

```{r, echo=FALSE, warning= FALSE}
  clin.data %>%
    group_by(institution.bin,ClusterID.Haralick) %>%
        summarize_each(funs(
    min_,
    Q1_,
    median_,
    Q3_,
    max_,
    sum_), sites)
```

Volume 


```{r, echo=FALSE, warning= FALSE}
  clin.data %>%
    group_by(institution.bin,ClusterID.Haralick) %>%
        summarize_each(funs(
    min_,
    Q1_,
    median_,
    Q3_,
    max_,
    sum_), volume.cc)
```

Number of cases

```{r, echo=FALSE, warning=FALSE}
  
  clin.data %>%
    group_by(institution.bin, ClusterID.Haralick) %>%
      do(data.frame(nrow=nrow(.)))
```

PFS 

```{r, echo=FALSE, warning=FALSE}

  clin.data%>%
    group_by(institution.bin,ClusterID.Haralick) %>%
      summarize_each(funs(
        min_,
        Q1_,
        median_,
        Q3_,
        max_
      ), pfs.mos)
```

OS 

```{r, echo=FALSE, warning=FALSE}

  clin.data%>%
    group_by(institution.bin,ClusterID.Haralick) %>%
      summarize_each(funs(
        min_,
        Q1_,
        median_,
        Q3_,
        max_
      ), os.mos)
```



Resection status

```{r, echo=FALSE, warning=FALSE}
  
  tdata <- clin.data%>%
    group_by(institution.bin,ClusterID.Haralick,residual.disease) %>%
       do(data.frame(nrow=nrow(.)))
  tdata <- na.omit(tdata)
  print(na.omit(tdata),n=50)
```

Were clusters associated with resection status

```{r, echo=FALSE, warning=FALSE}
  
  print(t(matrix(as.matrix(tdata[1:9,4]), nrow=3, ncol=3, dimnames = list(c("Complete", "Optimal", "Suboptimal"), c("High", "Low", "Medium")))))

  tidy(fisher.test(t(matrix(as.matrix(tdata[9:18,4]), nrow=3, ncol=3, dimnames = list(c("Complete", "Optimal", "Suboptimal"), c("High", "Low", "Medium")))), alternative = "two.sided"))

```

Fisher exact test for clusters vs. resection: TCIA

```{r, echo=FALSE, warning=FALSE}

  print(t(matrix(as.matrix(tdata[7:13,4]), nrow=3, ncol=2, dimnames = list(c("Complete", "Optimal", "Suboptimal"), c("High", "Low")))))

  tidy(fisher.test(t(matrix(as.matrix(tdata[7:13,4]), nrow=3, ncol=2, dimnames = list(c("Complete", "Optimal", "Suboptimal"), c("High", "Low")))), alternative = "two.sided"))


```


#### Prevalence of Platinum resistance in clusters

Platinum resistant

```{r, echo=FALSE, warning=FALSE}

  t <- clin.data%>%
    group_by(institution.bin,ClusterID.Haralick,platinum.resistant) %>%
       do(data.frame(nrow=nrow(.)))
  t <- na.omit(t)
  print(na.omit(t),n=21)
```


Fisher exact test for MSKCC clusters Haralick vs. platinum resistance

```{r, echo=FALSE, warning=FALSE}

  tidy(fisher.test(t(matrix(as.matrix.data.frame(t[1:6,4]), 
                            nrow=2,ncol=3, 
                            dimnames = list(c("Sensitive", "Resistant"), c("High", "Low", "Medium")))), 
                   alternative="two.sided"))

```

Fisher exact test for TCIA clusters Haralick vs. platinum resistance
```{r, echo=FALSE, warning=FALSE}
    tidy(fisher.test(t(matrix(as.matrix.data.frame(t[7:13,4]), 
                            nrow=2,ncol=3, 
                            dimnames = list(c("Sensitive", "Resistant"), c("High", "Low", "Medium")))), 
                   alternative="two.sided"))
```

***

##Differences in the texture values for scan manufacturers

```{r, echo=FALSE, warning=FALSE}
  clin.data$scanner <- clin.data$manufacturer=="GE"
  ge <- clin.data$scanner
  clin.data$scanner[ge] <- paste("GE")
  clin.data$scanner[!ge] <- paste("not GE")
  
```

cSE vs. Scanner

```{r, echo=FALSE, warning=FALSE}
  
  clin.data%>%
    do(tidy(wilcox.test(c.se ~ scanner, data = clin.data[clin.data$institution.bin!="MSKCC",])))
```


cluDev vs. scanner

```{r, echo=FALSE, warning=FALSE}

  clin.data%>%
    do(tidy(wilcox.test(clu.dev ~ scanner, data = clin.data[clin.data$institution.bin!="MSKCC",])))
```

cluDiss vs. scanner

```{r, echo=FALSE, warning=FALSE}

  clin.data%>%
    do(tidy(wilcox.test(clu.diss ~ scanner, data = clin.data[clin.data$institution.bin!="MSKCC",])))
```

Energy vs. scanner

```{r, echo=FALSE, warning=FALSE}

  clin.data%>%
    do(tidy(wilcox.test(energy ~ scanner, data = clin.data[clin.data$institution.bin!="MSKCC",])))
```

Contrast vs. scanner

```{r, echo=FALSE, warning=FALSE}

  clin.data%>%
    do(tidy(wilcox.test(contrast ~ scanner, data = clin.data[clin.data$institution.bin!="MSKCC",])))
```

Homogeneity vs. scanner

```{r, echo=FALSE, warning=FALSE}

  clin.data%>%
    do(tidy(wilcox.test(homogeneity ~ scanner, data = clin.data[clin.data$institution.bin!="MSKCC",])))
```

Entropy vs. scanner

```{r, echo=FALSE, warning=FALSE}

  clin.data%>%
    do(tidy(wilcox.test(entropy ~ scanner, data = clin.data[clin.data$institution.bin!="MSKCC",])))
```


## Generate the KM curves for using the ISTH clusters (low,mid vs. high, ultra-high)

PFS

```{r, echo=FALSE, warning=FALSE}
  #clin.data %>%
  #  group_by(institution.bin) %>%
     ggsurvplot(survfit(Surv(time=pfs.mos, event=recurrence=="1")~clusterISTH, data=clin.data[clin.data$institution.bin=="MSKCC",]), data=clin.data[clin.data$institution.bin=="MSKCC",], risk.table = TRUE, conf.int=TRUE, pval= TRUE, xlab="months", ylab="Proportion not progressed")

     ggsurvplot(survfit(Surv(time=pfs.mos, event=recurrence=="1")~clusterISTH, data=clin.data[clin.data$institution.bin!="MSKCC",]), data=clin.data[clin.data$institution.bin!="MSKCC",], risk.table = TRUE, conf.int=TRUE, pval= TRUE, xlab="months", ylab="Proportion not progressed")

 
```

OS


```{r, echo=FALSE, warning=FALSE}
#  clin.data %>%
#    group_by(institution.bin) %>%
     ggsurvplot(survfit(Surv(time=os.mos, event=dead=="1")~clusterISTH, data=clin.data[clin.data$institution.bin=="MSKCC",]), data=clin.data[clin.data$institution.bin=="MSKCC",], risk.table = TRUE, conf.int=TRUE, pval= TRUE, xlab="months", ylab="Proportion surviving")
 

  ggsurvplot(survfit(Surv(time=os.mos, event=dead=="1")~clusterISTH, data=clin.data[clin.data$institution.bin!="MSKCC",]), data=clin.data[clin.data$institution.bin!="MSKCC",], risk.table = TRUE, conf.int=TRUE, pval= TRUE, xlab="months", ylab="Proportion surviving")
```




